---
title: "CHM210 Assignment 1"
author: "YOUR NAME HERE"
date: '2022-XX-XX'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This is assignment 1, written in R Markdown. R Markdown is a formatting syntax that allows for easily combining R code with text in a word, html, or pdf document. R is a programming language commonly used for scientific computing for data analysis and visualization.The University of Toronto chemistry department has prepared an R textbook that you may find useful if you get stuck or confused, available here: https://uoftchem-teaching.github.io/R4EnvChem/.

In this assignment, we will be using R to import ozone, temperature, and pressure profiles with altitude, and do some data visualization and analysis.

Typically, the simplest method of importing data into R is by importing from a csv or other similarly formatted file. The function for doing so is the read.csv() function, as shown in the code block below.

```{r}
data <- read.csv("Assignment 1 Ozone.csv", sep = ",")
```

Once your data is imported, you can manipulate it either with the GUI, or using simple commands in R. For example:
